### 2021-05-26 analysis/word_lists.Rmd
add annotated word lists

---
title: "word lists"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(writexl)
library(ggthemes)
library(ggpubr)
```

# Exploring word lists

## Brysbaert & Biemiller 2017

```{r}
bb <-  read_xlsx("../data/word_lists/Brysbaert&Biemiller2017.xlsx")
```

> Brysbaert, M., & Biemiller, A. (2017). Test-based age-of-acquisition norms for 44 thousand English word meanings. Behavior research methods, 49(4), 1520-1523.

Variables:

 $ WORD        : from Dale and O’Rourke (1981)
 $ MEANING     : from Dale and O’Rourke (1981)
 $ AoAtestbased: updated test-based AoA measures from Dale and O’Rourke (1981) - values are in grades
 $ AoArating   : from Kuperman et al. (2012) - values are in years
 $ LWV         : original AoA measures from Dale and O’Rourke (1981) - values are in grades
 $ CDI         : age at which word is learned (produced according to CDI) - values are in month
 $ Morr        : age at which children could name pictures according to Morrison et al. (1997) - values are in month
 $ Floc        :suggested grade to teach the word according to the Flocabulary website - values are in grades

 ### Explore Morisson data

```{r}
mor <- bb %>%
  filter(!is.na(Morr))%>%
  filter(!duplicated(WORD))%>%
  mutate(aoa = round(Morr/12,2))%>%
  rename(english = WORD, 
         aoa_mor = aoa,
         aoa_rating_english = AoArating)%>%
  mutate_if(is.character, tolower)

```

```{r}
mor %>%
  group_by(aoa_mor)%>%
  summarise(n = n())%>%
  ggplot(aes(x = aoa_mor, y = n))+
  geom_bar(stat = "identity", alpha = .5, col = "black", fill = "firebrick")+
  theme_minimal()

mor%>%
  mutate(english = factor(english),
         english = fct_reorder(english, -aoa_mor))%>%
  ggplot(aes(y = english, x = aoa_mor, label = english))+
  geom_text(alpha = .5, size = 2)+
  facet_wrap(~aoa_mor, scales = "free", nrow = 1)+
  theme_few()+
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())

```
#### Relation to AOA ratings

```{r}
mor %>%
  ggplot(aes(x = aoa_rating_english, y = aoa_mor))+
  geom_abline(intercept = 0, slope = 1, lty = 2, col = "grey")+
  geom_point(alpha = 0.2)+
  theme_minimal()+
  stat_cor(method = "pearson")+
  xlim(0,12)+
  ylim(0,12)
```
## Schröder et al 2012

 > Schröder, A., Gemballa, T., Ruppin, S., & Wartenburger, I. (2012). German norms for semantic typicality, age of acquisition, and concept familiarity. Behavior research methods, 44(2), 380-394.

```{r}
setal <-  read_xls("../data/word_lists/schroederEtAl2012.xls")%>%
  mutate_if(is.character, tolower)%>%
  rename(aoa_german_s_m = aoa_m,
         aoa_german_s_sd = aoa_sd, 
         sem_cat = `semantic category`)

```
Variables:

 $ german                  : word in German
 $ english                 : British English translation
 $ semantic category       : semantic category of word (e.g. animals, birds, fruits, etc)
 $ exemplar_generation_n   : number of participants listing that item as exemplar for the semantic category it belongs to
 $ exemplar_generation_prop: prop of participants listing that item as exemplar for the semantic category it belongs to
 $ typicality_m            : mean for rating go of how typical is the item for the category (1-7 rating)
 $ typicality_sd           : sd for rating go of how typical is the item for the category (1-7 rating)
 $ aoa_german_s_m          : age of acquisition rating mean
 $ aoa_german_s_sd         : age of acquisition rating sd
 $ familiarity_m           : mean rating of degree to which they thought about or came in contact with a concept
 $ familiarity_sd          : sd rating of degree to which they thought about or came in contact with a concept
 $ word_freq               : normalized lemma frequency per million
 $ word_freq_log           : normalized lemma frequency per million in log scale
 $ word_length_phonemes    : number of phonemes
 $ word_length_syllables   : number of syllables

## Birchenough et al 2017

>Birchenough, J. M., Davies, R., & Connelly, V. (2017). Rated age-of-acquisition norms for over 3,200 German words. Behavior research methods, 49(2), 484-501.

```{r}
betal <-  read_csv("../data/word_lists/BirchenoughEtAL2017.csv", locale = locale(encoding = 'latin1'))%>%
  select(-Upper, AM_TWord)%>%
  mutate_if(is.character, tolower)%>%
  rename(german = Word, 
         english = BE_TWord, 
         aoa_german_b_m = AoAestimate,
         aoa_german_b_sd = SD)%>%
  select(german,english,RatperWord,aoa_german_b_m,aoa_german_b_sd)
```
 $ german         : word in German
 $ english        : British English translation
 $ RatperWord     : raters per word
 $ aoa_german_b_m : age of acquisition rating mean
 $ aoa_german_b_sd: age of acquisition rating sd

## Łuniewska et al 2019

> Łuniewska, M., Wodniecka, Z., Miller, C. A., Smolík, F., Butcher, M., Chondrogianni, V., ... & Haman, E. (2019). Age of acquisition of 299 words in seven languages: American English, Czech, Gaelic, Lebanese Arabic, Malay, Persian and Western Armenian. PloS one, 14(8), e0220611.

Total of 32 languages in the dataset, including German.

```{r}
letal <- read_xlsx("../data/word_lists/LuniewskaEtAl2019_aoa.xlsx")%>%
  mutate_if(is.character, tolower)%>%
  rename(aoa_german_l_m = German, 
         english = word)%>%
  select(english,aoa_german_l_m) 
```

## CLT dataset

```{r}
clt <-  read_delim("../data/word_lists/clt.csv", delim = ";")%>%
  mutate_if(is.character, tolower)%>%
  mutate(german = as.character(german))
```

## Merging Datasets

```{r}
md <- betal %>%
  select(-RatperWord, -aoa_german_b_sd)%>%
  full_join(setal%>%  select(german, english,word_freq,familiarity_m, aoa_german_s_m))%>%
  left_join(mor%>%select(english, aoa_mor, aoa_rating_english))%>%
  full_join(letal%>%left_join(clt))%>%
  filter(!is.na(german))%>%
  mutate(aoa_german_comb = ifelse(!is.na(aoa_german_b_m), aoa_german_b_m, 
                           ifelse(!is.na(aoa_german_s_m), aoa_german_s_m, aoa_german_l_m)))%>%
  select(german, word_freq,aoa_german_comb,clt,english,aoa_mor)%>%
  arrange(aoa_german_comb)%>%
  mutate(select = ifelse(clt == "yes", "x", ""), 
         word_freq = as.numeric(word_freq))%>%
  select(german, select, aoa_german_comb, word_freq, aoa_mor, clt, english)%>%
  mutate_if(is.numeric, round, digits = 2)
```

### Distribution of words by age

```{r}
md%>%
  ggplot(aes(x = aoa_german_comb))+
  geom_histogram(alpha = .5, fill = "firebrick",col = "black", position = "stack", binwidth = 0.25)+
  theme_minimal()+
  xlim(0,16)+
  scale_fill_viridis_d()
```
### Write Excel files for annotation

```{r}
md %>%
  filter(aoa_german_comb < 10)%>%
  mutate(group = as.numeric(factor(as.character(cut(row_number(), 10)))))%>%
  group_by(group)%>%
  group_walk(~write_xlsx(., paste0("../data/word_list_",.y$group,".xlsx")))

write_xlsx(md, path = "../data/word_list.xlsx")
```

### Relation between german AOA ratings

```{r}
md %>%
  ggplot(aes(x = aoa_german_l_m, aoa_german_b_m))+
  geom_abline(intercept = 0, slope = 1, lty = 2, col = "grey")+
  geom_point(alpha = 0.2)+
  theme_minimal()+
  geom_smooth(method = "lm", col = "firebrick")+
  stat_cor(method = "pearson")+
  xlim(0,12)+
  ylim(0,12)
```

### Relation between german and english AOA ratings

```{r}
md %>%
  select(german, aoa_german_b_m, aoa_german_s_m,aoa_german_l_m, aoa_mor)%>%
  pivot_longer(names_to = "source_german", values_to = "aoa_rating_german", cols = c(aoa_german_b_m, aoa_german_s_m,aoa_german_l_m))%>%
  ggplot(aes(x = aoa_rating_german, y = aoa_mor))+
  geom_abline(intercept = 0, slope = 1, lty = 2, col = "grey")+
  geom_point(alpha = 0.2)+
  theme_few()+
  facet_grid(~source_german)+
  geom_smooth(method = "lm", col = "firebrick")+
  stat_cor(inherit.aes = T, method = "pearson")+
  xlim(0,12)+
  ylim(0,12)
```

### Relation between german AOA rating and english test based AOA

```{r}
md %>%
  ggplot(aes(x = aoa_rating_german, aoa_mor))+
  geom_abline(intercept = 0, slope = 1, lty = 2, col = "grey")+
  geom_point(alpha = 0.2)+
  theme_minimal()+
  geom_smooth(method = "lm", col = "firebrick")+
  stat_cor(method = "pearson")+
  xlim(0,12)+
  ylim(0,12)
```

# Selected words from combined lists

For selections criteria see: `../docs/conding.md`.

```{r}
file.list <- list.files(path = "../data/annotated_word_lists/", pattern='*.xlsx', full.names = T)

ad <- lapply(file.list, read_xlsx)%>%bind_rows()%>%
  arrange(aoa_german_comb)%>%
  filter(select == "x")%>%
  select(1:7)

# add word type inforamtion

wt <- read_xlsx("../data/annotated_word_lists/word_list_collection_all.xlsx", skip = 1)%>%
  slice(-1)%>%
  pivot_longer(names_to = "word_type", values_to = "word", cols = 1:30)%>%
  mutate(word_type = gsub("\\..*","",word_type))%>%
  filter(!is.na(word))%>%
  mutate(word_type = recode(word_type, adj = "adjective", nouns = "noun", verbs = "verb"))%>%
  rename(german = word)%>%
  select(german, word_type)

fw <- ad%>%left_join(wt)%>%
  mutate(word_type = ifelse(german == "bürste", "noun", word_type),
         word_type = ifelse(german == "magnetisch", "adjective", word_type),
         word_type = ifelse(german == "gerechtigkeit", "noun", word_type))

write_csv(fw, file = "../data/word_list.csv")

```
## Distribution of selected words by age and word type

```{r}
fw%>%
  ggplot(aes(x = aoa_german_comb, fill = word_type))+
  geom_histogram(alpha = .5, col = "black", binwidth = 0.25, position = "identity")+
  theme_minimal()+
  xlim(1,11)+
  facet_grid(~word_type)+
  scale_fill_viridis_d()
```