### 2022-03-30 analysis/clt_distractors.Rmd

---
title: "CLT distractors"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(writexl)
library(googledrive)

```

```{r}
# drive_download("https://docs.google.com/spreadsheets/d/1xH3EOFpU9AQ7z9MtwQpnkjhbJOm8KyqV9SvZzdpezhw/edit#gid=0",
#                path = "./data/clt_new_words.xlsx",
#                overwrite = T)
```

```{r}

dist_select<- read_xlsx("./data/clt_new_words.xlsx")%>%
  filter(select == "x")%>%
  select(german)%>%
  mutate(unrelated = NA,
         semantic = NA,
         phonologic = NA)

write_xlsx(dist_select, "../data/clt_distractor_selection.xlsx")

```