### 2022-04-06 analysis/clt_data_processing.Rmd
more data

---
title: "CLT Results"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(jsonlite)
library(stringr)
library(ggpubr)
library(tidyboot)
library(readxl)
library(brms)
library(ggthemes)
library(writexl)
library(ggdist)
library(sf)
library(geomtextpath)
```

```{r}
# MPI
# rsync -zaP ccp-odcJ:/srv/ccp-odc/clt-extended/data/ ~/Work/Local/Quantex/vocab/data/clt_raw_data/

# ZAS und kischawi

# rsync -zaP ccp-odcJ:/srv/ccp-odc/clt-zas/data/ ~/Work/Local/Quantex/vocab/data/clt_raw_data_zas/
```

# Data

## MPI

```{r, warning=F}
files <- str_subset(dir("../data/clt_raw_data/", pattern = ".json"), pattern = "DELETE", negate = TRUE)

pilot <- str_subset(dir("../data/clt_pilot_data/", pattern = ".json"), pattern = "DELETE", negate = TRUE)

files <- files[!files %in% pilot]

exlcude <- dir("../data/clt_raw_data/", pattern = "DELETE")%>%
  as_tibble()%>%
  separate(value, into = c("rest", "subjID"), sep = "-")%>%
  mutate(subjID = str_replace(subjID, "DELETE",""))%>%
  select(subjID)

raw_data <- data_frame()
for (f in files) {
  jf <- paste("../data/clt_raw_data/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  subjID <- jd$meta$subjID
  order <- jd$meta$order
  add <- as_data_frame(jd$data) %>% mutate(subjID = subjID,
                                                 order = order)
  raw_data <- bind_rows(raw_data, add)
}

# filter out DELETE subjID
filtered_data <- raw_data%>%
  filter(!subjID %in% exlcude$subjID)%>%
  mutate(test_day = substr(timestamp, 0,10),
         test_day = as.Date(test_day,"%Y-%m-%d"))

# merge with aoa data
aoa <-  read_xlsx("../data/word_list.xlsx")%>%
  select(german,english, aoa_german_comb)%>%
  filter(german != "rakete",
         german != "hobel")%>%
  rename(targetWord = english)%>%
  mutate(targetWord = ifelse(targetWord == "frying pan", "fryingpan", targetWord),
         targetWord = ifelse(targetWord == "courgette", "zucchini", targetWord),
         targetWord = ifelse(targetWord == "rocket", "arugula", targetWord),
         targetWord = ifelse(targetWord == "palm tree", "palmtree", targetWord),
         targetWord = ifelse(targetWord == "airplane", "plane", targetWord))%>%
  filter(german != "kappe", 
         german != "karotte", 
         german != "mohrrübe")%>%
  distinct(german, .keep_all = T)%>%
  mutate(group = cut(aoa_german_comb, 
                   breaks=c(0, 3.5, 6, Inf), 
                   labels=c("low","middle","high")))

```

```{r}
dem <- bind_rows(
  read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 1),
  read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 2),
  read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 3),
  read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 4),
  read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 5),
  read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 7),
  read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 10),
  read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 12)
  )%>%
  select(code,g,`geb.tag`)%>%
  rename(subjID = "code", birth_day = "geb.tag", sex = "g")%>%
  mutate(birth_day = as.Date(birth_day, "%d.%m.%Y"),
         sex = ifelse(sex == "w","f","m"))%>%
  add_row(subjID = "leon", sex = "m", birth_day = as.Date("2015-09-03","%Y-%m-%d"))%>%
  add_row(subjID = "joshi", sex = "m", birth_day = as.Date("2017-06-01","%Y-%m-%d"))%>%
  distinct(subjID,.keep_all = T)

```

```{r}
dem%>%
  group_by(subjID)%>%
  summarise(n=n())%>%
  filter(n >1)

```

```{r}
data_mpi <- filtered_data%>%
  group_by(subjID)%>%
  distinct(trial, .keep_all = T)%>%
  mutate(correct = ifelse(targetWord == chosenWord, 1,0))%>%
  mutate(targetWord = ifelse(targetWord == "kangoroo", "kangaroo", targetWord))%>%
  left_join(aoa)%>%
  left_join(dem)%>%
  select(-timestamp)%>%
  ungroup()%>%
  mutate(age = as.numeric(difftime(test_day, birth_day, "days"))/365.25,
         age_group = substr(age,0,1))%>%
  select(-birth_day, -test_day)%>%
  filter(!is.na(age))%>%
  ungroup()

# filtered_data%>%
#   filter(trial == 1)%>%
#   distinct(subjID, .keep_all = T)%>%
#   left_join(dem)%>%
#   select(subjID, sex, birth_day, test_day)%>%
#   write_xlsx("../data/clt_participation_by_date.xlsx")
```

## ZAS and kischawi

```{r, warning=F}
# files_zas <- str_subset(dir("../data/clt_raw_data_zas/", pattern = ".json"), pattern = "DELETE", negate = TRUE) 
# 
# files_zas <- files_zas[!files_zas %in% pilot]
# 
# raw_data_zas <- data_frame()
# for (f in files_zas) {
#   jf <- paste("../data/clt_raw_data_zas/",f,sep="")
#   jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
#   subjID <- jd$meta$subjID
#   order <- jd$meta$order
#   add <- as_data_frame(jd$data) %>% mutate(subjID = subjID,
#                                                  order = order)
#   raw_data_zas <- bind_rows(raw_data_zas, add)
# }
# 
# proc_data_zas <- raw_data_zas%>%
#   mutate(subjID2 = subjID)%>%
#   separate(subjID2, into = c("bday","sex","postcode"), sep="_")%>%
#   mutate(test_day = substr(timestamp, 0,10),
#          test_day = as.Date(test_day,"%Y-%m-%d"),
#          bday  = as.Date(bday,"%d-%m-%Y"),
#          age = as.numeric(difftime(test_day, bday, units = "days"))/365.25,
#          sex = tolower(sex))
# 
# data_zas <- proc_data_zas%>%
#   left_join(aoa)%>%
#   select(-timestamp)%>%
#   ungroup()%>%
#   mutate(age_group = substr(age,0,1))%>%
#   select(-bday, -test_day)%>%
#   filter(!is.na(age))%>%
#   ungroup()

```

## Merge data

```{r}
data <- bind_rows(
  data_mpi%>%mutate(source = "mpi", postcode = "04103"),
  #data_zas%>%mutate(source = "zas_kischawi")
)
```

```{r}
write_csv(data,"../data/clt_clean_data.csv")
```

# Quality checks

```{r}
data%>%
  group_by(subjID)%>%
  summarise(n = n())

data%>%
  group_by(age_group)%>%
  summarise(n = length(unique(subjID)))

data%>%
  summarise(n = length(unique(subjID)))

data%>%
  group_by(source)%>%
  summarise(n = length(unique(subjID)))

data%>%
  filter(source == "mpi")%>%
  summarise(n = length(unique(subjID)))

filtered_data%>%
  summarise(n = length(unique(subjID)))

data%>%
  group_by(sex)%>%
  summarise(n = length(unique(subjID)))

```
# Reminder

```{r}
# reminder1 <- read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 15)
# 
# bind_rows(
#   read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 1),
#   read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 2),
#   read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 3),
#   read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 4),
#   read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 5),
#   read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 7),
#   read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 10),
#   read_xlsx("../../../../Cloud/Shared/CLT_Kinderlisten_fuer_Emailversandt_14032022.xlsx", sheet = 12)
#   )%>%
#   filter(!code %in% data$subjID)%>%
#   filter(!code %in% reminder1$code)%>%
#   filter(code != "RY5sK3xF")%>%
#   filter(code != "s6fdepZJ")%>%
#   write_xlsx("../data/clt_reminder2.xlsx")
```

# Descriptives

```{r}
data%>%
  group_by(subjID)%>%
  summarise(age = round(mean(age),2))%>%
  ggplot(aes(x = age))+
  geom_jitter(aes(y = 0),width =0, height = 0.02, pch = 1, alpha =.75)+
  geom_density()+
  #ylim(-1,1)+
  theme_minimal()

```

```{r}
p1 <- data%>%
  group_by(targetWord, aoa_german_comb, german)%>%
  tidyboot_mean(col = correct)

ggplot(p1, aes(x = aoa_german_comb, y = mean))+
  stat_cor(aes(label = paste(..r.label..)),r.accuracy = 0.01, cor.coef.name = "r", label.y = 0.85)+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), alpha = .75)+
  geom_textpath(aes(label = german), hjust = -.2) +
  #geom_point( alpha = .75)+
  geom_smooth(method = "lm")+
  labs(x = "Rated age of acquisition", y = "Proportion correct")+
  ylim(0,1)+
  theme_minimal()
```

```{r}
p1%>%
  mutate(targetWord = factor(targetWord))%>%
  ggplot(., aes(x = reorder(targetWord, -aoa_german_comb), y = mean))+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), alpha = .75)+
  #geom_point( alpha = .75)+
  labs(x = "Item", y = "Proportion correct")+
  ylim(0,1)+
  theme_minimal()+
  coord_flip()
```

```{r}
p1_1 <- data%>%
  group_by(targetWord,german, order)%>%
  tidyboot_mean(col = correct)

p1_2 <- p1_1 %>%
  select(-empirical_stat, -n)%>%
  pivot_wider(values_from = c(mean, ci_lower,ci_upper), names_from = "order")

ggplot(p1_2, aes(x = mean_orderA, y = mean_orderB))+
  stat_cor(aes(label = paste(..r.label..)),r.accuracy = 0.01, cor.coef.name = "r")+
  geom_abline(intercept = 0, slope = 1, lty = 3, alpha = .75)+
  geom_textpath(aes(label = german), hjust = -.2) +
  #geom_hline(yintercept = 0.25, alpha = .5)+
  #geom_vline(xintercept = 0.25, alpha = .5)+
  geom_errorbar(aes(xmin = ci_lower_orderA, xmax = ci_upper_orderA), alpha = .5)+
  geom_errorbar(aes(ymin = ci_lower_orderB, ymax = ci_upper_orderB), alpha = .5)+
  geom_point( size =2, stroke = 1, alpha= .75)+
  labs(x = "Order A", y = "Order B")+
  ylim(0,1)+
  xlim(0,1)+
  theme_minimal()
```

```{r}
p2 <- data%>%
  group_by(subjID, age)%>%
  tidyboot_mean(col = correct)

ggplot(p2, aes(x = age, y = mean))+
  #stat_cor()+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), alpha = .25)+
  #geom_point( alpha = .75)+
  geom_smooth(method = "lm")+
  labs(x = "Age", y = "Proportion correct")+
  ylim(0,1)+
  xlim(3,8.2)+
  theme_minimal()
```

```{r}
dist <- data%>%
  filter(correct ==0)%>%
  group_by(targetWord, aoa_german_comb, chosenCategory)%>%
  summarise(n = n())

ggplot(dist, aes(x = reorder(targetWord,aoa_german_comb), y = n, fill = chosenCategory))+
  geom_bar(stat ="identity")+
  scale_fill_colorblind(name = "Distractor type", labels = c("Semantic","Phonological","Unrelated"))+
  theme_minimal()+
  labs(x = "Target word", y = "Count")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
                legend.direction = "vertical",
        legend.position = c(0.1,0.8), 
        legend.spacing.y = unit(0, 'cm'))
```

# Location

```{r}
library(mapproj)
library(raster)
```

```{r}
germany <- getData(country = "Germany", level = 2)%>%
  st_as_sf()

postcodes <- read_csv("../data/plz-5stellig-daten.csv")%>%
  separate(note, into = c("zip","NAME_2"), sep=" ")
```