### 2022-04-19 analysis/clt_gafo_merge.Rmd

---
title: "CLT Gafo merge"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(tidyboot)
library(brms)
library(tidybayes)
```

```{r}
clt_data <- read_csv("../data/clt_clean_data.csv")

gafo_data <- readRDS("../data/gafo-testtrials.rds")%>%
  ungroup()%>%
  mutate(testday = as.Date(substr(timestamp,0,10), "%Y-%m-%d"), 
         dob = as.character(dob),
         dob = ifelse(dob == "0016-05-01","2016-05-01",dob),
         dob = ifelse(dob == "0018-08-04","2018-08-04",dob),
         dob = as.Date(dob,"%Y-%m-%d"))
```

```{r}
merge <- clt_data%>%
  group_by(age_group,age,subjID)%>%
  summarise(mean_clt = mean(correct, na.rm = T), 
            cor_clt = sum(correct, na.rm = T),
            n_clt = length(correct))%>%
  inner_join(
    gafo_data%>%
      mutate(gafo_age = as.numeric(difftime(testday, dob, "days"))/365.25)%>%
      group_by(gafo_age,subjID)%>%
      summarise(mean_gaze = mean(correctBox, na.rm = T))
  )%>%
  mutate(time_diff = age - gafo_age)%>%
  distinct(subjID, .keep_all = T)

```

```{r}
merge%>%
  ungroup()%>%
  #group_by(age_group)%>%
  summarise(n = length(unique(subjID)))

merge%>%
  ungroup()%>%
  group_by(age_group)%>%
  summarise(n = length(unique(subjID)))
```

```{r}
ggarrange(
  ggplot(merge, aes(x = time_diff))+
  geom_histogram(alpha = .75, col = "black", fill = "white", bins = 100)+
  labs(x = "Time (in years) between tests", y = "Count")+
  scale_color_viridis_c(name = "Age")+
  #ylim(0.7,1)+
  #coord_fixed(3)+
  theme_minimal(),

  ggplot(merge, aes(x = age, y = mean_clt))+
  geom_jitter( width = 0.005, height = 0.005, alpha = .75)+
  stat_cor(cor.coef.name = "r")+
  geom_smooth(method = "lm", col = "firebrick")+
  labs(x = "Age (in years)", y = "Vocabulary score")+
  #scale_color_viridis_c(name = "Age")+
  #ylim(0.7,1)+
  #coord_fixed(3)+
  theme_minimal(),

  ggplot(merge, aes(x = mean_gaze, y = mean_clt, col = age))+
  geom_jitter( width = 0.005, height = 0.005, alpha = .75)+
  stat_cor(cor.coef.name = "r")+
  geom_smooth(method = "lm", col = "firebrick")+
  labs(x = "Gaze understanding score", y = "Vocabulary score")+
  scale_color_viridis_c(name = "Age")+
  #ylim(0.7,1)+
  #coord_fixed(3)+
  theme_minimal(),
  widths = c(1,1),
  common.legend = T,
  nrow = 1
)
```

```{r}
merge_model <- merge%>%
  ungroup()%>%
  mutate(age = scale(age),
         mean_gaze = scale(mean_gaze))
```

```{r}
mfull <- brm( cor_clt|trials(n_clt)~age+mean_gaze,
     data = merge_model,
     family = binomial
     )%>%
  add_criterion(c("loo","waic"))

mage <- brm( cor_clt|trials(n_clt)~age,
     data = merge_model,
     family = binomial
     )%>%
  add_criterion(c("loo","waic"))

mgaze <- brm( cor_clt|trials(n_clt)~mean_gaze,
     data = merge_model,
     family = binomial
     )%>%
  add_criterion(c("loo","waic"))

mnull <- brm( cor_clt|trials(n_clt)~1,
     data = merge_model,
     family = binomial
     )%>%
  add_criterion(c("loo","waic"))
```

```{r}
loo_compare(mfull, mage, mgaze,mnull)%>%as_tibble(rownames = "model")
```

```{r}
plot(mfull)

pp_check(mfull)

mfull
```

```{r}
p1 <- mfull %>%
  gather_draws(b_age, b_mean_gaze)%>%
  mutate(Predictor = recode(.variable,
                        b_age = "Age (scaled)",
                        b_mean_gaze = "Gaze understanding score (scaled)"))

ggplot(p1,aes(y = Predictor, x = .value, fill = stat(x) > 0)) +
  stat_halfeye(alpha = .7, .width = c(.95, .80)) +
    labs(x = "Model Estimate", y ="")+
  guides(fill = "none")+
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c("grey","#85AD00"))+
  xlim(-0.2, 0.5)+
  theme_minimal()
```