### 2022-04-21 analysis/clt_item_selection.Rmd

---
title: "CLT item selection"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(ggthemes)
```

```{r}
data <- read_csv("../data/clt_clean_data.csv")

model_params <- read_csv("../data/model_parameters_3PL.csv")
```

# Correlation subtest and full test

```{r}
select <-model_params %>%
  pivot_wider(names_from = nlpar, values_from = Estimate)%>%
  arrange(
    Easiness, 
    #-Discrimination
    )%>%
  slice(1:15)%>%
  pull("item")

full <- data%>%
  group_by(age_group,subjID)%>%
  summarise(mean_full = mean(correct))

sub <- data%>%
  filter(targetWord %in% select)%>%
  group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))

comp <- full%>%left_join(sub)

ggplot(comp, aes(x = mean_full, y = mean_sub))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_jitter(pch = 1, alpha = .75, height = 0.01, width = 0.01)+
  stat_cor()+
  facet_wrap(~age_group)+
  ylim(0,1.02)+
  xlim(0,1.02)+
  theme_few()

```

```{r}
dat <- tibble(
  iter = c(),
  items = c(),
  cor = c(),
)

full <- data%>%
  group_by(subjID)%>%
  summarise(mean_full = mean(correct))

for (i in 1:200) {

  items <-params %>% pull("item")

  sel_items <- sample(items, 15)

  sub <- data%>%
  filter(targetWord %in% sel_items)%>%
  group_by(subjID)%>%
  summarise(mean_sub = mean(correct))

  cor <- cor(full$mean_full,sub$mean_sub)

  df <- tibble(
    iter = i,
    items = sel_items,
    cor = cor
  )

  dat <- bind_rows(dat,df)

}

dat%>%
  group_by(items)%>%
  summarise(mean_cor = mean(cor))%>%
  arrange(-mean_cor)

full <- data%>%
  group_by(age_group,subjID)%>%
  summarise(mean_full = mean(correct))

select2 <- dat%>%group_by(items)%>%
  summarise(mean_cor = mean(cor))%>%
  arrange(-mean_cor)%>%slice(1:15)%>%pull(items)

sub2 <- data%>%
  filter(targetWord %in% select2)%>%
  group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))

comp2<- full%>%left_join(sub2)

ggplot(comp2, aes(x = mean_full, y = mean_sub))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_jitter(pch = 1, alpha = .75, height = 0.01, width = 0.01)+
  stat_cor()+
  facet_wrap(~age_group)+
  ylim(0,1.02)+
  xlim(0,1.02)+
  theme_few()
```

# Optimum subset selection via simulated annealing

```{r}

score_fn <- function(subset) {
	easinesses <- sort(easiness[subset])
	nn_dists <- rep(0, sum(subset)-1)
	for(i in 1:sum(subset)-1) {
		nn_dists[i] <- easinesses[i+1] - easinesses[i]
	}
	spacing <- -1*sd(nn_dists)
	mean_discrim <- mean(discrimination[subset])
	return(spacing + mean_discrim / 6)
}

proposal_fn <- function(subset) {
	swaps <- sample(c(1, 1, 1, 2, 2, 3), 1)
	for(i in 1:swaps) {
		active_items <- seq(1:length(subset))[subset == TRUE]
		inactive_items <- seq(1:length(subset))[subset == FALSE]
		active_to_swap <- sample(active_items, 1)
		inactive_to_swap <- sample(inactive_items, 1)
		subset[active_to_swap] <- FALSE
		subset[inactive_to_swap] <- TRUE
	}
	return(subset)
}

simulated_annealing <- function(k) {
	easiness <- filter(model_params, nlpar=="Easiness") %>% pull(Estimate)
	discrimination <- filter(model_params, nlpar=="Discrimination") %>% pull(Estimate)
	N <- length(easiness)

	best_subset <- sample(c(rep(TRUE, k), rep(FALSE, N-k)))
	best_score <- score_fn(best_subset)

	temp <- 100
	rejected <- 0
	for(i in 1:1e6) {
		new_subset <- proposal_fn(best_subset)
		new_score <- score_fn(new_subset)
		accept_decrease <- rbernoulli(1, temp / 100)
		if(new_score > best_score | accept_decrease) {
			best_subset <- new_subset
			best_score <- new_score
			#print(best_score)
		} else {
			rejected <- rejected + 1
		}
		if(rejected == 100000) {
			break
		}
		temp <- temp*0.999
	}
	print(best_score)
	return(best_subset)
}

sim_a <- simulated_annealing(15)

items <- model_params%>%filter(nlpar=="Easiness")
sel_items <- items$item[sim_a == TRUE]

```

## Visualize parameters of selected items

```{r}
model_params%>%
  filter(item %in% sel_items)%>%
  ggplot(aes(reorder(item, -aoa_german_comb), Estimate, ymin = Q2.5, ymax = Q97.5)) +
	facet_wrap("nlpar", scales = "free_x") +
	geom_pointrange() +
	coord_flip() +
	labs(x = "Item")+
  theme_minimal()
```
## Correaltion with full test

```{r}
full <- data%>%
  group_by(age_group,subjID)%>%
  summarise(mean_full = mean(correct))

sub2 <- data%>%
  filter(targetWord %in% sel_items)%>%
  group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))

comp2<- full%>%left_join(sub2)

ggplot(comp2, aes(x = mean_full, y = mean_sub))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_jitter(pch = 1, alpha = .75, height = 0.01, width = 0.01)+
  stat_cor()+
  facet_wrap(~age_group)+
  ylim(0,1.02)+
  xlim(0,1.02)+
  theme_few()
```