### 2022-04-25 analysis/clt_item_selection.Rmd
push

---
title: "CLT item selection"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(ggthemes)
```

```{r}
data <- read_csv("../data/clt_clean_data.csv")%>%
  filter(age_group != 8)

all_model_params <- read_csv("../data/model_parameters_3PL.csv")
```

# Optimum subset selection via simulated annealing

```{r}
# exclude items based on differential item functioning

exclude <- c("choir", "flounder", "badger")

model_params <- all_model_params%>%
  filter(!item %in% exclude)
```

```{r}
easb <- tibble()

for (j in 1:100) {

for (i in 1:48) {

  x <- sample(easiness, i)

	nn_dists <- rep(0, length(x)-1)
	for(k in 1:length(x)-1) {
		nn_dists[k] <- x[k+1] - x[k]
	}

	spacing <- -1*sd(nn_dists)

	row <- tibble(iter = j,
                size = i, 
                mean = spacing/2)
  easb <- bind_rows(easb,row)
}
}

ggplot(easb, aes(x = factor(size), y = mean))+
  geom_point(pch = 1)+
  theme_minimal()

```

```{r}
discrimb <- tibble()

for (j in 1:100) {

for (i in 1:48) {
  x <- sample(discrimination, i)
  mean_x <- mean(x)
  row <- tibble(iter = j,
                size = i, 
                mean = mean_x/2)
  discrimb <- bind_rows(discrimb,row)
}
}

ggplot(discrimb, aes(x = factor(size), y = mean))+
  geom_point(pch = 1)+
  theme_minimal()

```

```{r}
spreadb <- tibble()

for (j in 1:100) {

for (i in 1:48) {
  x <- sample(easy_spread, i)
  mean_x <- mean(x)
  row <- tibble(iter = j,
                size = i, 
                mean = mean_x/5)
  spreadb <- bind_rows(spreadb,row)
}
}

ggplot(spreadb, aes(x = factor(size), y = mean))+
  geom_point(pch = 1)+
  theme_minimal()

```

```{r}
items <- model_params%>%filter(nlpar=="Easiness")%>%pull(item)
easiness <- filter(model_params, nlpar=="Easiness") %>% pull(Estimate)
easy_spread <- model_params%>%filter(nlpar=="Easiness")%>%mutate(width = abs(Q97.5 - Q2.5)) %>% pull(width)
discrimination <- filter(model_params, nlpar=="Discrimination") %>% pull(Estimate)

full <- data%>%
  group_by(subjID)%>%
  summarise(mean_full = mean(correct))

score_fn <- function(subset) {
	easinesses <- sort(easiness[subset])
	nn_dists <- rep(0, sum(subset)-1)
	for(i in 1:sum(subset)-1) {
		nn_dists[i] <- easinesses[i+1] - easinesses[i]
	}
	spacing <- -1*sd(nn_dists)
	mean_discrim <- mean(discrimination[subset])
	mean_spread <- mean(easy_spread[subset])

	# sel <- items[subset == TRUE]
	# 
	# sub_dat <- data%>%
	#   filter(targetWord %in% sel)%>%
	#   group_by(subjID)%>%
	#   summarise(mean_sub = mean(correct))
	# 
	# cor <- cor(sub_dat$mean_sub,full$mean_full)
	# 
	# return(spacing + mean_discrim/6 + cor/3)
	return(spacing + mean_discrim/2 - mean_spread/5)
}

proposal_fn <- function(subset) {
	swaps <- sample(c(1, 1, 1, 2, 2, 3), 1)
	for(i in 1:swaps) {
		active_items <- seq(1:length(subset))[subset == TRUE]
		inactive_items <- seq(1:length(subset))[subset == FALSE]
		active_to_swap <- sample(active_items, 1)
		inactive_to_swap <- sample(inactive_items, 1)
		subset[active_to_swap] <- FALSE
		subset[inactive_to_swap] <- TRUE
	}
	return(subset)
}

simulated_annealing <- function(k) {
  easiness <- filter(model_params, nlpar=="Easiness") %>% pull(Estimate)
  easy_spread <- model_params%>%filter(nlpar=="Easiness")%>%mutate(width = abs(Q97.5 - Q2.5)) %>% pull(width)
  discrimination <- filter(model_params, nlpar=="Discrimination") %>% pull(Estimate)
	N <- length(easiness)

	best_subset <- sample(c(rep(TRUE, k), rep(FALSE, N-k)))
	best_score <- score_fn(best_subset)

	temp <- 100
	rejected <- 0
	for(i in 1:1e6) {
		new_subset <- proposal_fn(best_subset)
		new_score <- score_fn(new_subset)
		accept_decrease <- rbernoulli(1, temp / 100)
		if(new_score > best_score | accept_decrease) {
			best_subset <- new_subset
			best_score <- new_score
			#print(best_score)
		} else {
			rejected <- rejected + 1
		}
		if(rejected == 10000) {
			break
		}
		temp <- temp*0.999
	}
	#print(best_score)
	ret <- tibble(best_subset = list(best_subset), 
	              best_score = best_score)

	return(ret)
}
```

```{r, message=F, warning=F, comment=F}
sim_a <- simulated_annealing(20)

sel_items <- items[unlist(sim_a$best_subset) == TRUE]

sim_a$best_score
```

## Visualize parameters of selected items

```{r}
model_params%>%
  filter(item %in% sel_items)%>%
  ggplot(aes(reorder(item, -Estimate), Estimate, ymin = Q2.5, ymax = Q97.5)) +
	facet_wrap(~nlpar, scales = "free_x") +
	geom_pointrange() +
	coord_flip() +
	labs(x = "Item")+
  theme_minimal()
```

## Correaltion with full test

```{r}
sub2 <- data%>%
  filter(targetWord %in% sel_items)%>%
  group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))

comp2<- full%>%left_join(sub2)

ggplot(comp2, aes(x = mean_full, y = mean_sub))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_jitter(pch = 1, alpha = .75, height = 0.01, width = 0.01)+
  stat_cor()+
  facet_wrap(~age_group)+
  ylim(0,1.02)+
  xlim(0,1.02)+
  theme_few()
```

## Variation in score

```{r, warning=FALSE}
sim_score <- tibble()

for (i in 5:35) {

  for (j in 1:10) {

   sim <- simulated_annealing(i)

   score <- tibble(score = sim$best_score, 
                   size = i,
                   iter = j)

   sim_score <- bind_rows(sim_score, score)
  }

}

sim_score%>%
  ggplot(aes(x = size, y = score, group = iter))+
  geom_point(pch = 1)+
  theme_minimal()
```

## Determine size

```{r, warning=FALSE}
sim_size <- tibble()

for (j in 1:5) {

for (i in 5:45) {

   sim <- simulated_annealing(i)

    sel <- items[unlist(sim$best_subset) == TRUE]

    cor <- data%>%
      filter(targetWord %in% sel)%>%
      group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))%>%
  right_join(full)%>%
  group_by(age_group)%>%
  summarise(cor = cor(mean_sub, mean_full))%>%
  mutate(size = i, 
         iter = j)

  sim_size <- bind_rows(sim_size, cor)
}

}

sim_size%>%
  ggplot(aes(x = factor(size), y = cor, col = factor(age_group)))+
  geom_point(pch = 1)+
  geom_line(aes(group = interaction(age_group,factor(iter)), lty = factor(iter)))+
  theme_minimal()+
  #facet_grid(factor(iter)~.)+
  labs(x = "No. of items", y = "Correlation with full test")+
  scale_color_colorblind(name = "Age group")

```

## Select items

```{r}
item_sel <- tibble()

for (i in 1:100) {

   sim <- simulated_annealing(20)

   sel <- items[unlist(sim$best_subset) == TRUE]

   it <- data%>%
     distinct(targetWord)%>%
     filter(targetWord %in% sel)%>%
     select(targetWord)%>%
     mutate(iter = i)

   item_sel <- bind_rows(item_sel, it)
}

item_sel%>%
  group_by(targetWord)%>%
  summarise(n = n()/max(iter))%>%
  ggplot(aes(x = reorder(targetWord, -n), y = n))+
  geom_bar(stat = "identity", col = "black", fill = "white")+
  labs(x = "Item", y = "Proportion selected")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## ICC for subset

```{r}
frq_sel_items <- item_sel%>%
  group_by(targetWord)%>%
  summarise(n = n()/max(iter))%>%
  arrange(-n)%>%
  slice(1:20)%>%
  pull(targetWord)

icc3 %>% 
  filter(item %in% frq_sel_items)%>%
  ggplot(aes(x = theta, y = p,group = item, col = aoa_german_comb)) +
  #geom_line() +
  geom_textline(aes(label = item)) +
  #geom_texthline(yintercept = 0.25, lty = 3, alpha = .75, label = "guessing",hjust = 0.8)+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  #facet_wrap(~group)+
  #guides(col = F)+
  scale_color_viridis_c(name = "AoA") +
  labs(title = "ICCs for selected items",
       x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  ylim(0,1)+
  theme_minimal()
```
## Correaltion selected with full test

```{r}
subsel <- data%>%
  filter(targetWord %in% frq_sel_items)%>%
  group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))

compsel<- full%>%left_join(subsel)

ggplot(compsel, aes(x = mean_full, y = mean_sub))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_jitter(pch = 1, alpha = .75, height = 0.01, width = 0.01)+
  stat_cor()+
  facet_wrap(~age_group)+
  ylim(0,1.02)+
  xlim(0,1.02)+
  theme_few()
```

```{r}
model_params%>%
  filter(item %in% frq_sel_items)%>%
  ggplot(aes(reorder(item, -Estimate), Estimate, ymin = Q2.5, ymax = Q97.5)) +
	facet_wrap(~nlpar, scales = "free_x") +
	geom_pointrange() +
	coord_flip() +
	labs(x = "Item")+
  theme_minimal()
```