### 2022-05-16 analysis/clt_visuals.Rmd
add model params

---
title: "CLT extension visuals"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(ggthemes)
```

```{r}
data <- read_csv("../data/clt_clean_data.csv")%>%
  filter(age_group != 8)

all_model_params <- read_csv("../data/model_parameters_3PL.csv")

params_sex <- readRDS("../saves/model_params_irt3_dif_sex.rds")%>%
  select(-uci, -lci)%>%
  pivot_wider(names_from = sex, values_from = mode)%>%
  mutate(sex_dif = f-m)

icc3 <- readRDS("../saves/icc_3PL.rds")
```

```{r}
all_model_params%>%
  filter(nlpar == "Easiness")%>%
  ggplot(aes(x = Estimate, y = aoa_german_comb))+
  #geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_point()+
  stat_cor()
```