### 2022-08-11 analysis/visuals.Rmd
clean repo

---
title: "oREV visuals"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(ggthemes)
library(tidyboot)
library(readxl)
library(geomtextpath)
library(ggridges)
library(scales)
```

```{r}
data <- read_csv("../data/clean_data.csv")

all_model_params <- read_csv("../data/model_parameters_3PL.csv")

params_sex <- readRDS("../saves/model_params_irt3_dif_sex.rds")%>%
  select(-uci, -lci)%>%
  pivot_wider(names_from = sex, values_from = mode)%>%
  mutate(sex_dif = f-m)

icc3 <- readRDS("../saves/icc_3PL.rds")

frq_sel_items <- readRDS("../saves/selected_items.rds")

sim_size <- readRDS("../saves/sim_size.rds")

sim_size_age <- readRDS("../saves/sim_size_age.rds")

params_joy <- readRDS("../saves/model_params_irt3_draws.rds")
```

# Data descriptives

## Figure 2

```{r}
p2 <- data%>%
  group_by(source,subjID, age)%>%
  tidyboot_mean(col = correct)

page <- ggplot(p2, aes(x = age, y = mean))+
  #stat_cor()+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  #geom_point(alpha = .25)+
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), alpha = .5, size = .25)+
  #geom_point( alpha = .75)+
  geom_smooth(method = "lm", col = "firebrick", alpha = .25)+
  labs(x = "Participant age", y = "Proportion correct (by participant)")+
  ylim(0,1)+
  xlim(3,8)+
  theme_few()
```

```{r}
p1 <- data%>%
  group_by(targetWord, source, aoa_german_comb, german)%>%
  tidyboot_mean(col = correct)

paoa <- p1%>%
ggplot(aes(x = aoa_german_comb, y = mean, col = source))+
  stat_cor(aes(x = aoa_german_comb, y = mean,label = paste(..r.label..)),inherit.aes = F,r.accuracy = 0.01, cor.coef.name = "r",label.y = 0.85)+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), alpha = .75, pch = 1)+
  #facet_grid(~measure)+
  #geom_textpath(aes(label = german), hjust = -.2) +
  #geom_point( alpha = .75)+
  geom_smooth(method = "lm", col = "firebrick", alpha = .25)+
  scale_color_colorblind(name = "")+
  labs(x = "Age of acquisition (adult rated)", y = "Proportion correct (by item)")+
  ylim(0,1)+
  theme_few()+
  theme(legend.position = c(0.2,0.5))
```

```{r}
sex_dif_id <- data%>%
  group_by(subjID,sex)%>%
  summarise(mean = mean(correct))

order_dif_id <- data%>%
  group_by(subjID,order)%>%
  summarise(mean = mean(correct))

dif_id <- bind_rows(
  sex_dif_id%>%rename(comparison = sex)%>%mutate(type = "by sex"),
  order_dif_id%>%rename(comparison = order)%>%mutate(type = "by order")
)%>%
  mutate(comparison = recode(comparison,
                             m = "Male",
                             f = "Female",
                             orderA = "Order A", 
                             orderB = "Order B",))

sex_dif <- data%>%
  group_by(sex)%>%
  tidyboot_mean(col = correct)

order_dif <- data%>%
  group_by(order)%>%
  tidyboot_mean(col = correct)

dif <- bind_rows(
  sex_dif%>%rename(comparison = sex)%>%mutate(type = "by sex"),
  order_dif%>%rename(comparison = order)%>%mutate(type = "by order")
)%>%
  mutate(comparison = recode(comparison,
                             m = "Male",
                             f = "Female",
                             orderA = "Order A", 
                             orderB = "Order B",))

```

```{r}
pdif <- ggplot(dif, aes(x = comparison, y = mean))+
  geom_dotplot(data = dif_id, aes(x = comparison, y = mean), binaxis='y', stackdir='center',
               stackratio=1.5, dotsize=.2, alpha = .2)+
  geom_bar(stat = "identity", alpha = .25)+
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), alpha = .75)+
  facet_grid(~type, scales = "free_x" )+
  scale_y_continuous(limits=c(0.5,1),oob = rescale_none)+
  labs(y = "Proportion correct (by participant)", x = "")+
  theme_few()
```

```{r}
ggarrange(page, paoa,pdif, nrow = 1, labels = c("A","B", "C"), widths = c(1,1,1))
```

```{r}
ggsave("../graphs/data_fig.png", width = 12, height = 3, scale = 1.25, bg = "white")
```

## Figure 3

```{r}
params_sex_p <- params_sex%>%
  mutate(nlpar = factor(nlpar, levels = c("Easiness", "Discrimination"), labels = c("Difficulty", "Discrimination")))%>%
  group_by(nlpar)%>%
  mutate(sd = sd(sex_dif), 
         cut_low = mean(sex_dif)-2*sd,
         cut_high = mean(sex_dif)+2*sd)%>%
  mutate(ex = ifelse(sex_dif >cut_high | sex_dif < cut_low, "out", "in"))

ggplot(params_sex_p, aes(x = sex_dif, label = item, y = reorder(item, abs(sex_dif)), col = ex))+
  geom_vline(aes(xintercept = cut_low), lty = 3, alpha = .75)+
  geom_vline(aes(xintercept = cut_high), lty = 3, alpha = .75)+
  geom_point(alpha = .5)+
  scale_color_manual(values = c("black", "firebrick"))+
  guides(col = "none")+
  labs(x = "Difference between estimates (female - male)", y = "Target word")+
  facet_grid(~nlpar)+
  theme_few()
```
```{r}
ggsave("../graphs/dif_fig.png", width = 4, height = 6, scale = 1.25, bg = "white")
```

## Figure 4

```{r}
picc <- icc3 %>% 
  mutate(select = ifelse(item %in% frq_sel_items,"yes", "no"),
         pos = as.numeric(as.factor(item)))

iccp <- ggplot(picc%>%filter(select == "yes"), aes(x = theta, y = p,group = item)) +
  #geom_line(alpha = .5) +
  geom_line(data = picc%>%filter(select == "no"), aes(x = theta, y = p,group = item), col = "grey", alpha = .5)+
  #geom_line(aes(x = theta, y = p,group = item, col = item))+
  geom_textline(aes(label = item, col = item),  gap = T, size = 3) +
  #geom_texthline(yintercept = 0.25, lty = 3, alpha = .75, label = "guessing",hjust = 0.8)+
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  #facet_wrap(~group)+
  guides(col = F)+
  scale_color_viridis_d()+
  labs(#title = "ICCs for selected items",
       x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  ylim(0,1)+
  theme_few()
```

```{r}
pss <- sim_size%>%
  group_by(size)%>%
  summarise(mean_cor = mean(cor),
         min = min(cor),
         max = max(cor))

pssa <- sim_size_age%>%
  group_by(size, age_group)%>%
  summarise(mean_cor = mean(cor),
         min = min(cor),
         max = max(cor))

psel <- ggplot()+
  geom_vline(xintercept = 20, alpha = .5, col = "#31493C")+
  geom_pointrange(data = pssa,aes(x = size, y = mean_cor,ymin=min, max = max, col = factor(age_group)),pch = 1, alpha = .5)+
  geom_line(data = pssa, aes(group = factor(age_group), x = size, y = mean_cor, col = factor(age_group)), alpha = .5)+
  geom_pointrange(data = pss,aes(x = size, y = mean_cor,ymin=min, max = max),pch = 1)+
  geom_line(data = pss, aes(group = 1, x = size, y = mean_cor))+
  scale_linetype(name = "Run")+
  scale_x_continuous(breaks = c(5,10,15,20,25,30,35,40))+
  scale_color_ptol(name = "Age group")+
  theme_few()+
  ylim(0.6, 1)+
  theme(legend.position = c(.8,.4))+
  labs(x = "No. of items in subset", y = "Correlation with full task")
```

```{r}
pparam <- all_model_params%>%
  filter(item %in% frq_sel_items)%>%
  mutate(Estimate = ifelse(nlpar == "Easiness", -Estimate, Estimate),
         Q2.5 = ifelse(nlpar == "Easiness", -Q2.5, Q2.5),
         Q97.5 = ifelse(nlpar == "Easiness", -Q97.5, Q97.5),
         nlpar = factor(nlpar, levels = c("Easiness", "Discrimination"), labels = c("Difficulty", "Discrimination")))%>%
  ggplot(aes(reorder(item, Estimate), col = item, Estimate, ymin = Q2.5, ymax = Q97.5)) +
	facet_wrap(~nlpar, scales = "free_x") +
  scale_color_viridis_d()+
	geom_pointrange(pch = 1) +
	coord_flip() +
  guides(col = F)+
	labs(x = "Target word")+
  theme_few()

# pparam2 <- params_joy%>%
#   left_join(all_model_params)%>%
#   filter(item %in% frq_sel_items)%>%
#     mutate(Estimate = ifelse(nlpar == "Easiness", -Estimate, Estimate),
#          nlpar = factor(nlpar, levels = c("Easiness", "Discrimination"), labels = c("Difficulty", "Discrimination")))%>%
#   ggplot( aes(y = reorder(item, Estimate), x = val, fill = item))+
#   geom_density_ridges(size = 0.2,scale = 4, alpha = .5)+
#   facet_grid(~nlpar, scales = "free_x")+
#   labs(y = "Target word", x = "Estimate")+
#   scale_fill_viridis_d()+
#   guides(fill = F)+
#   theme_minimal()
```

```{r}
# ggarrange(
#           ggarrange(psel,pparam2, ncol = 1, labels = c("A","B"), heights = c(1,1.2)),iccp,
#           widths = c(1,1.5), labels = c("","C"))

ggarrange(
          ggarrange(psel,pparam, ncol = 1, labels = c("A","B"), heights = c(1,1)),iccp,
          widths = c(1,1.5), labels = c("","C"))
```

```{r}
ggsave("../graphs/item_fig2.png", width = 12, height = 5, scale = 1.25, bg = "white")
```