### 2022-11-23 analysis/orev_item_selection.Rmd

---
title: "oREV item selection and analysis"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(ggthemes)
library(tidybayes)
library(brms)
library(rstan)
library(loo)
library(coda)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}

func <- function(x){
  abs(1-x)
}
```

```{r}
data <- read_csv("../data/clean_data.csv")

irt_dat <- data%>%
  select(subjID, targetWord, correct, order, sex, aoa_german_comb)

aoa <- data%>%distinct(targetWord, .keep_all = T)%>%select(targetWord, aoa_german_comb)

full <- data%>%
  group_by(subjID)%>%
  summarise(mean_full = mean(correct))
```

# Models

## Rasch Model

```{r}
prior_1pl_guess <- 
  prior("normal(0, 2)", class = "b", nlpar = "eta") +
  prior("normal(0, 1)", class = "sd", group = "subjID", nlpar = "eta") + 
  prior("normal(0, 3)", class = "sd", group = "targetWord", nlpar = "eta")
```

### Model

```{r}
irt1_guess <- brm(
  data = irt_dat,
  family = brmsfamily("bernoulli", "identity"),
  bf(
    correct ~ 0.25 + 0.75 * inv_logit(eta),
    eta ~ 1 + (1 | targetWord) + (1 | subjID),
    nl = TRUE
  ),
  prior = prior_1pl_guess,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  cores = 6,
  chains = 6,
  iter = 6000,
  threads = threading(8), #to speed things up, comment out if not on a cluster
  backend = "cmdstanr" #to speed things up, comment out if not on a cluster
)%>%add_criterion(c("loo","waic"))

saveRDS(irt1_guess, "../saves/irt1_guess.rds")

irt1_guess <- readRDS("../saves/irt1_guess.rds")
```

### Model to extract fit indices

```{r}
prior_1pl_guess <- prior("normal(0, 2)", class = "b", nlpar = "eta") +
  prior("normal(0, 1)", class = "sd", group = "subjID", nlpar = "eta") +
  prior("normal(0, 3)", class = "sd", group = "targetWord", nlpar = "eta")

irt1_guess_standata <- brms:::make_standata(
  formula = bf(
    correct ~ 0.25 + 0.75 * inv_logit(eta),
    eta ~ 1 + (1 | targetWord) + (1 | subjID),
    nl = TRUE
  ),
  data = irt_dat,
  family = brmsfamily("bernoulli", "identity"),
  prior = prior_1pl_guess
)

irt1_guess_itemloo <- stan(
  file = "../analysis/irt1_guess_itemloo.stan",
  data = irt1_guess_standata[c("N", "Y", "N_1", "J_1", "N_2", "J_2")],
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  cores = 6,
  chains = 6,
  iter = 6000
)

saveRDS(irt1_guess_itemloo, "../saves/irt1_guess_itemloo.rds")

irt1_guess_itemloo <- readRDS("../saves/irt1_guess_itemloo.rds")

irt1_guess_itemloo_loom <- loo(irt1_guess_itemloo, pars = "loglik_m")

saveRDS(irt1_guess_itemloo_loom, file = "../saves/irt1_guess_looitems_loom.rds")

irt1_guess_itemloo_idxs <- irt_dat %>%
  mutate(item_id = as.integer(factor(targetWord))) %>%
  group_by(targetWord) %>%
  summarize(
    mean_correct = mean(correct),
    item_id = unique(item_id),
    targetWord = unique(targetWord)
  ) %>%
  full_join(mutate(as_tibble(irt1_guess_itemloo_loom$pointwise[, "influence_pareto_k", drop = FALSE]), item_id = row_number())) %>%
  rename(pareto_k = influence_pareto_k) %>%
  full_join(
    as_draws_df(irt1_guess_itemloo) %>% 
      select(.chain, .iteration, starts_with("infit"), starts_with("outfit"), starts_with("r_2")) %>% 
      pivot_longer(-c(.chain, .iteration), names_pattern = "(.+)\\[([0-9]+)\\]", names_to = c(".value", "item_id")) %>% 
      rename(threshold = r_2_eta_1) %>% 
      mutate(item_id = as.integer(item_id)) %>% 
      group_by(item_id) %>% 
      summarize(across(c(infit, outfit, threshold), list(mean = mean, mode = estimate_mode, hdi_lower = hdi_lower, hdi_upper = hdi_upper)))
  ) %>%
  arrange(pareto_k) %>%
  print(n = nrow(.))

saveRDS(irt1_guess_itemloo_idxs, file = "../saves/irt1_guess_itemloo_idxs.rds")
```

## 2PL Model

```{r}
prior_va_2pl <- 
  prior("normal(0, 2)", class = "b", nlpar = "eta") +
  prior("normal(0, 1)", class = "b", nlpar = "logalpha") +
  prior("normal(0, 1)", class = "sd", group = "subjID", nlpar = "eta") + 
  prior("normal(0, 3)", class = "sd", group = "targetWord", nlpar = "eta") +
  prior("normal(0, 1)", class = "sd", group = "targetWord", nlpar = "logalpha")
```

```{r}
irt2PL <- brm(
  data = irt_dat,
  family = brmsfamily("bernoulli", "identity"),
  bf(
    correct ~ 0.25 + 0.75 * inv_logit(exp(logalpha) * eta),
    eta ~ 1 + (1 |i| targetWord) + (1 | subjID),
    logalpha ~ 1 + (1 |i| targetWord),
    nl = TRUE
  ),
  prior = prior_va_2pl,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  cores = 6,
  chains = 6,
  iter = 6000,
  threads = threading(8), #to speed things up, comment out if not on a cluster
  backend = "cmdstanr" #to speed things up, comment out if not on a cluster
)%>%add_criterion(c("loo","waic"))

saveRDS(irt2PL, "../saves/irt2PL.rds")

irt2PL <- readRDS("../saves/irt2PL.rds")

coef2PL <- coef(irt2PL)
```

# Indices

```{r}
items <- irt1_guess_itemloo_idxs %>% arrange(targetWord)%>%pull(targetWord)
easiness_rasch <- irt1_guess_itemloo_idxs %>% arrange(targetWord)%>%pull(threshold_mode)
infit <- irt1_guess_itemloo_idxs %>% arrange(targetWord)%>% pull(infit_mode)
outfit <- irt1_guess_itemloo_idxs%>% arrange(targetWord) %>% pull(outfit_mode)
disc_2PL <- coef2PL$targetWord[, , "logalpha_Intercept"] %>% as_tibble(rownames = "item")%>%pull(Estimate)
pkd <- irt1_guess_itemloo_idxs%>% arrange(targetWord) %>% pull(pareto_k)
```

# Item selection

## Score

```{r}
score_dist <- tibble()

for (j in 1:16) {

for (i in 10:35) {

  x <- sample(easiness_rasch, i)

	nn_dists <- rep(0, length(x)-1)

	for(k in 1:length(x)-1) {
		nn_dists[k] <- x[k+1] - x[k]
	}

	x2 <- sample(disc_2PL, i)
  x_dist2 <- unlist(lapply(x2, func))

  x3 <- sample(infit, i)
  x_dist3 <- unlist(lapply(x3, func))

  x4 <- sample(outfit, i)
  x_dist4 <- unlist(lapply(x4, func))

  x <- sample(pkd, i)

  mean_spacing <- -1*sd(nn_dists)/3
  var_2PL <- -1*var(x_dist2)*10
  mean_infit <- -2*mean(x_dist3)
  mean_outfit <- -4*mean(x_dist4)
  mean_pkd <- mean(x)/5

  sum_score = mean_spacing + mean_infit + mean_outfit+ var_2PL +mean_pkd

	row <- tibble(iter = j,
                size = i, 
                spacing = mean_spacing,
	              infit = mean_infit,
	              outfit = mean_outfit,
	              var_2PL_disc = var_2PL, 
	              pkd = mean_pkd, 
	              score = sum_score)

  score_dist <- bind_rows(score_dist,row)
}
}

score_dist%>%
  pivot_longer(cols = c(-size, -iter), names_to = "type", values_to = "value")%>%
  ggplot(aes(x = factor(size), y = value, col = type))+
  geom_point(pch = 1)+
  facet_wrap(~iter)+
  ggtitle("Score = spacing + infit + outfit+ var_2PL_disc + pkd")+
  theme_bw()

```

```{r}
ggsave("../graphs/score.png", height = 12, width = 16, scale = 1)
```

## Simulated annealing algorithm

```{r}
score_fn <- function(subset) {
	easinesses <- sort(easiness_rasch[subset])
	nn_dists <- rep(0, sum(subset)-1)
	for(i in 1:sum(subset)-1) {
		nn_dists[i] <- easinesses[i+1] - easinesses[i]
	}
	spacing <- -1*sd(nn_dists)/3

	var_disc_sample <- disc_2PL[subset]
	var_disc_2PL <- -1*var(var_disc_sample)*10

	infit_sample <- infit[subset]
	infit_dist <- unlist(lapply(infit_sample, func))
  mean_infit <- -4*mean(infit_dist)

	outfit_sample <- outfit[subset]
	outfit_dist <- unlist(lapply(outfit_sample, func))
  mean_outfit <- -2*mean(outfit_dist)

  mean_pkd <- mean(pkd[subset])/5

	# sel <- items[subset == TRUE]
	# 
	# sub_dat <- data%>%
	#   filter(targetWord %in% sel)%>%
	#   group_by(subjID)%>%
	#   summarise(mean_sub = mean(correct))
	# 
	# cor <- cor(sub_dat$mean_sub,full$mean_full)
	# 
	# return(spacing + mean_discrim/6 + cor/3)
	return(spacing + mean_infit + mean_outfit+ var_disc_2PL +mean_pkd)
}

proposal_fn <- function(subset) {
	# Randomly sample a number of swaps.
	# Prefer a small number of swaps for "fine tuning", but allow
	# occasional large numbers of swaps, including a complete
	# exchange of the subset
	subset_size = sum(as.integer(subset))
	max_swaps = min(subset_size, length(subset) - subset_size)
	swaps <- rbinom(1, max_swaps-1, 1/(max_swaps-1)) + 1

	# Choose the items to swap
	active_items <- seq(1:length(subset))[subset == TRUE]
	inactive_items <- seq(1:length(subset))[subset == FALSE]
	actives_to_swap <- sample(active_items, swaps)
	inactives_to_swap <- sample(inactive_items, swaps)

	# Do the swapping
	for(i in 1:swaps) {
		subset[actives_to_swap[i]] <- FALSE
		subset[inactives_to_swap[i]] <- TRUE
	}
	return(subset)
}

simulated_annealing_rasch <- function(k, cooling_ratio=0.999, reset_thresh=1000, break_thresh=10000) {

  items <- irt1_guess_itemloo_idxs %>% arrange(targetWord)%>%pull(targetWord)
  easiness_rasch <- irt1_guess_itemloo_idxs %>% arrange(targetWord)%>%pull(threshold_mode)
  infit <- irt1_guess_itemloo_idxs %>% arrange(targetWord)%>% pull(infit_mode)
  outfit <- irt1_guess_itemloo_idxs%>% arrange(targetWord) %>% pull(outfit_mode)
  disc_2PL <- coef2PL$targetWord[, , "logalpha_Intercept"] %>% as_tibble(rownames = "item")%>%pull(Estimate)
  pkd <- irt1_guess_itemloo_idxs%>% arrange(targetWord) %>% pull(pareto_k)

  N <- length(easiness_rasch)

	current_subset <- sample(c(rep(TRUE, k), rep(FALSE, N-k)))
	best_subset <- current_subset
	best_score <- score_fn(best_subset)

	temp <- 100
	rejected <- 0
	no_new_bests <- 0
	for(i in 1:1e6) {
		# Score new subset, and toss a coin
		new_subset <- proposal_fn(current_subset)
		new_score <- score_fn(new_subset)
		accept_decrease <- rbernoulli(1, temp / 100)

		# Accept the new subset if it's an improvement, or if our
		# cooling coin came up heads.
		if(new_score > best_score | accept_decrease) {
			current_subset <- new_subset
			rejected <- 0
			if(new_score > best_score) {
				best_subset <- new_subset
				best_score <- new_score
				no_new_bests <- 0
			} else {
				no_new_bests <- no_new_bests + 1
			}
		# Quit if we've had too many rejections in a row.
		} else {
			rejected <- rejected + 1
			no_new_bests <- no_new_bests + 1
			if(rejected == break_thresh) {
				#print(best_score)
			  ret <- tibble(best_subset = list(best_subset),
	              best_score = best_score)

				return(ret)
			}
		}
		# Start random resets to the current best subset if we haven't
		# found anything better in quite a while.
		if(no_new_bests > reset_thresh & rbernoulli(1, 1/100)) {
			current_subset <- best_subset
		}

		# Cool it!
		temp <- temp*cooling_ratio
	}
	#print(best_score)
	ret <- tibble(best_subset = list(best_subset),
	              best_score = best_score)

	return(ret)
}
```

### Test run

```{r, message=F, warning=F, comment=F}
sim_rasch_test <- simulated_annealing_rasch(20)

sel_items_rasch_test <- items[unlist(sim_rasch_test$best_subset) == TRUE]

sub_rasch_test <- data%>%
  filter(targetWord %in% sel_items_rasch_test,
         )%>%
  group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))

comp_rasch_test<- full%>%left_join(sub_rasch_test)

ggplot(comp_rasch_test, aes(x = mean_full, y = mean_sub))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_jitter(pch = 1, alpha = .75, height = 0.01, width = 0.01)+
  stat_cor()+
  facet_wrap(~age_group)+
  ylim(0,1.02)+
  xlim(0,1.02)+
  theme_few()
```

## Check variation in score

```{r, warning=FALSE}
sim_score_rasch <- tibble()

for (i in 5:40) {

  for (j in 1:5) {

   sim <- simulated_annealing_rasch(i)

   score <- tibble(score = sim$best_score, 
                   size = i,
                   iter = j)

   sim_score_rasch <- bind_rows(sim_score_rasch, score)
  }

}

saveRDS(sim_score_rasch, "../saves/variation_in_score.rds")

sim_score_rasch%>%
  ggplot(aes(x = size, y = score, group = iter))+
  geom_point(pch = 1)+
  theme_minimal()
```

## Determine size

```{r}
irt_comp_size <- tibble()

  for(j in 2:2){

  for (i in 38:40) {

    sim <- simulated_annealing_rasch(i)

    sel <- items[unlist(sim$best_subset) == TRUE]

    sub_dat <- irt_dat%>%filter(targetWord %in% sel)

    m1PL <- update(irt1_guess, newdata =sub_dat, chains = 6, cores = 6, threads = threading(8), backend = "cmdstanr")%>%add_criterion(c("loo"))
  	m3PL <- update(irt2PL, newdata =sub_dat, chains = 6, cores = 6, threads = threading(8), backend = "cmdstanr")%>%add_criterion(c("loo"))

  	comp <- loo_compare(m1PL, m3PL)%>%as_tibble(rownames = "model")%>%filter(model == "m1PL")%>%mutate(ratio = abs(elpd_diff)/(2*se_diff))

  	sub_dat_cor <- sub_dat%>%
  	  group_by(subjID)%>%
  	  summarise(mean_sub = mean(correct))

  	cor <- cor(sub_dat_cor$mean_sub,full$mean_full)

  	row <- tibble(size = i, 
           iter = j,
           elpd_diff = as.numeric(comp%>%pull(elpd_diff)),
           se_diff = as.numeric(comp%>%pull(se_diff)),
           ratio = as.numeric(comp%>%pull(ratio)),
           correlation = cor)

    irt_comp_size <- bind_rows(irt_comp_size, row)

    saveRDS(irt_comp_size, "../saves/irt_comp_size.rds")
  }

  }

saveRDS(irt_comp_size, "../saves/irt_comp_size.rds")

irt_comp_size <- readRDS("../saves/irt_comp_size.rds")

irt_comp_size%>%
  mutate(ratio = ifelse(elpd_diff == 0,0,ratio))%>%
  pivot_longer(cols = c(ratio, correlation), names_to = "type",values_to = "value")%>% 
  ggplot(aes(x = factor(size), y = value))+
  geom_point(pch = 1)+
  geom_line()+
  facet_grid(type~. , scales = "free_y")+
  theme_bw()+
  labs(x = "No. of items", y = "Comparison value", title = "ratio = abs(elpd_diff)/(2*se_diff)")
```

```{r}
ggsave("../graphs/item_sel.png", height = 6, width = 10, scale = 1)
```

## Extract selected items

```{r}
item_sel_rasch <- tibble()

for (i in 1:10) {

   sim <- simulated_annealing_rasch(22)

   sel <- items[unlist(sim$best_subset) == TRUE]

   it <- data%>%
     distinct(targetWord)%>%
     filter(targetWord %in% sel)%>%
     select(targetWord)%>%
     mutate(iter = i)

   item_sel_rasch <- bind_rows(item_sel_rasch, it)
}

item_sel_rasch%>%
  group_by(targetWord)%>%
  summarise(n = n()/max(iter))%>%
  ggplot(aes(x = reorder(targetWord, -n), y = n))+
  geom_bar(stat = "identity", col = "black", fill = "white")+
  labs(x = "Item", y = "Proportion selected")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

rasch_selected_items <- item_sel_rasch%>%
  group_by(targetWord)%>%
  summarise(n = n()/max(iter))%>%
  arrange(-n)%>%
  pull(targetWord)

saveRDS(rasch_selected_items, "../saves/selected_items_rasch.rds")
```

# Rasch model with selected items

## Fit model

```{r}
irt_dat_sel <- irt_dat%>%
  filter(targetWord %in% rasch_selected_items)
```

```{r}
irt1_guess_sel <- brm(
  data = irt_dat_sel,
  family = brmsfamily("bernoulli", "identity"),
  bf(
    correct ~ 0.25 + 0.75 * inv_logit(eta),
    eta ~ 1 + (1 | targetWord) + (1 | subjID),
    nl = TRUE
  ),
  prior = prior_1pl_guess,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  cores = 6,
  chains = 6,
  iter = 6000,
  threads = threading(8), #to speed things up, comment out if not on a cluster
  backend = "cmdstanr" #to speed things up, comment out if not on a cluster
)%>%add_criterion(c("loo","waic"))

saveRDS(irt1_guess_sel, "../saves/irt1_guess_sel.rds")

irt1_guess_sel <- readRDS("../saves/irt1_guess_sel.rds")
```

## ICC

```{r}
icc1_guess_sel <- posterior_samples(irt1_guess_sel)%>% 
  select(b_eta_Intercept, starts_with("r_targetWord"))%>%
  mutate(iter = 1:n()) %>% 
  pivot_longer(starts_with("r_targetWord"), names_to = "item", values_to = "xi") %>%
  mutate(item = str_extract(string = item, pattern = "(?<=\\[).*(?=,Intercept\\])"))%>%
  expand(nesting(iter, b_eta_Intercept, item, xi),
         theta = seq(from = -6, to = 6, length.out = 100)) %>% 
  mutate(p = 0.25 + 0.75*inv_logit_scaled((b_eta_Intercept + theta + xi))) %>%  
  group_by(theta, item) %>% 
  summarise(p = mean(p))%>%
  left_join(aoa%>%rename(item = targetWord))
```

```{r}
icc1_guess_sel %>% 
  ggplot(aes(x = theta, y = p,group = item, col = aoa_german_comb)) +
  geom_line() +
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75)+
  scale_color_viridis_c(name = "AoA") +
  labs(title = "ICCs for the 1PL",
       x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  ylim(0,1)+
  theme_minimal()
```

## Differential Item Functioning

### Sex

```{r}
irt1_guess_dif_sex_sel <-  brm(
  data = irt_dat_sel,
  family = brmsfamily("bernoulli", "identity"),
  bf(
    correct ~ 0.25 + 0.75 * inv_logit(eta),
    eta ~ 1 + (0+ sex | targetWord) + (1 | subjID),
    nl = TRUE
  ),
  prior = prior_1pl_guess,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  cores = 6,
  chains = 6,
  iter = 6000
)%>%add_criterion(c("loo","waic"))

saveRDS(irt1_guess_dif_sex_sel, "../saves/irt1_guess_dif_sex_sel.rds")

irt1_guess_dif_sex_sel<- readRDS("../saves/irt1_guess_dif_sex_sel.rds")

```

### Order

```{r}
irt1_guess_dif_order_sel <- brm(
  data = irt_dat_sel,
  family = brmsfamily("bernoulli", "identity"),
  bf(
    correct ~ 0.25 + 0.75 * inv_logit(eta),
    eta ~ 1 + (0+ order | targetWord) + (1 | subjID),
    nl = TRUE
  ),
  prior = prior_1pl_guess,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  cores = 6,
  chains = 6,
  iter = 6000
)%>%add_criterion(c("loo","waic"))

saveRDS(irt1_guess_dif_order_sel, "../saves/irt1_guess_dif_order_sel.rds")

irt1_guess_dif_order_sel<- readRDS("../saves/irt1_guess_dif_order_sel.rds")

```

#### Comapre models

```{r}
loo_compare(irt1_guess_sel, irt1_guess_dif_order_sel, irt1_guess_dif_sex_sel)%>%as_tibble(rownames = "model")

brms::bayes_R2(irt1_guess)
```

#### Visual model inspection

```{r}
dif_sex_guess <- as_draws_df(irt1_guess_dif_sex_sel)%>%
  select(b_eta_Intercept, starts_with("r_targetWord"))%>%
  mutate(iter = 1:n()) %>%
  pivot_longer(starts_with("r_targetWord")) %>%
  mutate(name = str_remove(name, pattern = "r_targetWord__eta"),
         name = str_remove_all(name, pattern = "\\[|\\]"))%>%
  separate(name, into = c("item", "sex"), sep = "\\,")%>%
  mutate(sex = str_remove(sex, pattern = "sex"),
         val =  value)%>%
  group_by(item, sex)%>%
  summarise(mode = estimate_mode(val),
            uci = hdi_upper(val),
            lci = hdi_lower(val))%>%
  left_join(aoa%>%rename(item = targetWord))

saveRDS(dif_sex_guess, "../saves/model_params_irt1_guess_dif_sex.rds")

dif_sex_guess <- readRDS("../saves/model_params_irt1_guess_dif_sex.rds")

ggplot(dif_sex_guess,aes(x = reorder(item, aoa_german_comb))) +
	geom_point(aes(col = sex, y = lci), position = position_dodge(width = .5)) +
  geom_point(aes(col = sex, y = uci), position = position_dodge(width = .5)) +
  geom_linerange(aes(col = sex, ymin = lci + 0.1, ymax = uci-0.1), position = position_dodge(width = .5), alpha = .5) +
	coord_flip() +
  scale_color_colorblind(labels = c("male","female"), name = "Group")+
	labs(x = "Item", y = "Easiness estimate")+
  theme_few()
```

```{r}
ggsave("../graphs/dif_ci.png", height = 10, width = 10, scale = 1)
```

```{r}
dif_sex_guess%>%
  pivot_wider(names_from = sex, values_from = c(mode,uci,lci))%>%
  ggplot(., aes(x = mode_f, y = mode_m))+
  geom_abline(intercept = 0, slope = 1, lty = 1, alpha = .75)+
  geom_point(pch = 1, size = 2, stroke  = 1, aes(col = factor(aoa_german_comb)))+
  geom_linerange(aes(ymin = lci_m, ymax = uci_m),  alpha = .25, lty = 1)+
  geom_linerange(aes(xmin = lci_f, xmax = uci_f),  alpha = .25, lty = 1)+
  geom_text(aes(label = item, x = uci_f +0.2))+
  labs(x = "Group: female", y = "Group: male")+
  scale_color_viridis_d()+
  guides(col = F)+
  theme_few()

```

```{r}
ggsave("../graphs/dif_cor.png", height = 8, width = 8, scale = 1.2)
```

# Reliability

## KR20

```{r}
rel_dat <- irt_dat%>%
  select(-sex, -order)%>%
  filter(targetWord %in% rasch_selected_items)%>%
  group_by(subjID)%>%
  distinct(targetWord, .keep_all = T)%>%
  pivot_wider(names_from = targetWord, values_from = correct)%>%
  ungroup()%>%
  select(-subjID)

kr20_rel <- kr20(rel_dat, hit = 1)

kr20_rel
```

## Andrich Reliability

```{r}

pers_params <- ranef(irt1_guess_sel)$subjID%>%as_tibble(rownames = "subjID")

sep_rel <- 1 - 
  (1/length(pers_params$Estimate.eta_Intercept) * sum(pers_params$Est.Error.eta_Intercept^2))/
  (1/(length(pers_params$Estimate.eta_Intercept)-1)*sum((pers_params$Estimate.eta_Intercept - mean(pers_params$Estimate.eta_Intercept))^2))

sep_rel
```